Grouping and summarizing

Summarizing the median life expectancy

```{r}
gapminder %>%
  summarise(medianLifeExp = median(lifeExp))   # obtengo un resultado opera la funcion a lo largo del set de datos

# Filer for 1957 then summarize the median life expentancy
gapminder %>%
  filter(year == 1957) %>%
  summarise(medianLifeExp = median(lifeExp))


gapminder %>%
  filter(year == 1957) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```

Summarizing by year

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```

Summarizing by continent

```{r}
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))%>%
  arrange(desc(maxGdpPercap))

```

Summarizing by continent and year

```{r}
gapminder %>%
  group_by(continent, year) %>%              # Agrupo por medio de dos variables
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))


```
Line Plots

Visualizing median life expenctancy over time

```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_year

ggplot(by_year, aes(x=year, y=medianGdpPercap)) +
  geom_line()                 # grafica de lineas

```

```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x=year, y=medianGdpPercap, color=continent)) +
  geom_line()

```
Bar Plots

Visualizing median GDP per capita by continent

```{r}
by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_continent

ggplot(by_continent, aes (x=continent , y=medianGdpPercap)) +
  geom_col()                          # grafica de barras

```
Visualizing GDP per capita by country in Oceania

```{r}
oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year == 1952)

oceania_1952

ggplot(oceania_1952, aes(x=country, y=gdpPercap)) +
  geom_col()

```

Histogram

Visualizing population

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)

gapminder

gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil = pop/1000000)

#Creando la visualizacion
ggplot(gapminder_1952, aes(x=pop_by_mil)) +
  geom_histogram(bins=50)   # bins se utiliza para delimitar la cantidad de divisiones en el histograma
  
```

Visualizing population with x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop)) +
  geom_histogram(bins=50) +
  scale_x_log10()
  
```
Boxplots

Comparing GDP per capita across continents

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

gapminder_1952 %>%
  filter(continent=="Americas")

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() +
  scale_y_log10()

```

Adding a title to your graph

```{r}
ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() +
  ggtitle("Comparing GDP per capita across continents")

```

Data Manipulation with dplyr

```{r}

library(tidyverse)
counties <- read.csv('counties.csv')
glimpse(counties)   # glimpse ver de la perspectiva de la estructura del dataset
counties  # counties contiene el senso de Estados Unidos que se realizo el 2015

```

Selecting columns

```{r}
counties %>%
  select(state, county, population, men, women)

counties_selected <- counties %>%
  select(state,county,population,private_work,public_work,self_employed)

counties_selected %>%
  arrange(desc(public_work))   # Arrange se utiliza para realizar ordenamientos, en este caso es ordenamiento descendente

counties_selected <- counties %>%
  select(state, county, population)

counties_selected %>%
  filter(population > 1000000)

counties_selected <- counties %>%
  select(state,county,population,private_work,public_work,self_employed)

counties_selected %>%
  filter(state=="Texas", population > 10000) %>%
  arrange(desc(private_work))
  
```

Select, mutate, filter and arrange

```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))

```

The verb count

```{r}
select_counties <- counties %>%
  select(state, region, county, population, citizens)

counties_selected <- counties %>%
  count(region, sort = TRUE)  # sort ordena de forma ascendente

```

Counting citizens by state


```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)


counties_selected %>%
  count(state, wt=citizens) %>%       # wt (wait) agrupa por estado pero el conteo lo hace por la variable citizens
  arrange(desc(n))

```

Mutating and counting

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)

counties_selected %>%
  mutate(population_walk = population*walk/100) %>%
  count(state, wt=population_walk, sort=TRUE)

```

Summarizing

```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)

counties_selected %>%
  summarize(min_population=min(population), max_unemployment=max(unemployment), averagen_income=mean(income))

```

Summarizing by state

```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarize(total_area=sum(land_area), total_population=sum(population)) %>%
  mutate(density=total_population/total_area) %>%
  arrange(desc(density))

counties_selected <- counties %>%
  select(region, state, county, population)

counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop=sum(population)) %>%
  summarize(average_pop=mean(total_pop), median_pop=median(total_pop))

```

Top_n verb

```{r}
counties_selected <- counties %>%
  select(region, state, county, metro,population, walk)

counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)

counties_selected <- counties %>%
  select(region, state, county, population, income)

counties_selected %>%
  group_by(region, state) %>%
  summarize(average_income=mean(income)) %>%
  top_n(1, average_income)

```
Ungroup

```{r}
counties_selected <- counties %>%
  select(state, metro, population)

counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop=sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>%
  count(metro)

```
